---
title: "Cicindela_formosa_PATTERN"
author: "Wei Han"
date: "2023-08-19"
output: html_document
---


```{r}
library(tidyverse)
library(patternize)
library(dplyr)
library(recolorize)
library(raster)
library(palmerpenguins)
library(ggplot2)
library(cowplot)
```

[[[[PATTERNIZE OBJECTS]]]]

PATTERNIZE
```{r}

setwd("C:/Users/USER/Documents/My School Documents/UofA/Lightroom")

#ID list of images
pictures <- tools::file_path_sans_ext(dir("test images/", ".jpg"))

# make list with images
imageList <- makeList(pictures, type = "image", prepath = "test images/", extension = ".jpg")

# make list with landmarks
landmarkList <- makeList(pictures, type = "landmark", prepath = "test landmarks/", extension = "_landmarks.txt")

# use CNC1860122 as the (test) reference image
target <- landmarkList[["CNC1151335"]]

# make mask
mask <- read.table("test masks increased/CNC1151335 outline.txt", header = FALSE)

# Use patternize function alignLan() to align images
imageList_aligned <- alignLan(imageList, 
                              landmarkList, 
                              transformRef = target, 
                              adjustCoords = TRUE,
                              plotTransformed = T, 
                              resampleFactor = 5, 
                              cartoonID = 'CNC1151335',
                              maskOutline = mask)

```


[[[[PREPARING RECOLORIZE]]]]]

```{r}

setwd("C:/Users/USER/Documents/My School Documents/UofA/Lightroom")

# Reading back Raster file - still a Raster at this point
imageList_aligned <- readRDS("rds files/imageList_aligned_partII.rds")

# Changing rasters to image arrays 
imgs <- lapply(imageList_aligned, brick_to_array)

# save raster extents for later conversion:
extent_list1 <- lapply(imageList_aligned, extent)

# Using the same names as imageList_aligned to imgs (image array and not raster)
names(imgs) <- names(imageList_aligned)

# Seeing what the image arrays look like (why is it so squashed?)
plotImageArray(imgs[["CNC1151335"]])

```

[[[[[RECOLORIZE]]]]

```{r}

# make an empty vector (a one-dimensional list) for storing the recolorize objects
rc_list <- vector("list", length(imgs))

# use the same names as imgs to rc_list
names(rc_list) <- names(imgs)

# make a list of recolorize images. For every image, run the same recolorize2() function to fit a recolorize object (Bins = 3, which means that 27 bins will be formed in the colour space because dividing each colour channel into 3 bins forms a total of 27 bins (3^3) since there are 3 colour channels

for (i in 1:length(imgs)) {
  rc_list[[i]] <- recolorize2(imgs[[i]], bins = 3,
                              cutoff = 35, plotting = FALSE)
}

```
[[[[VISUALIZE I]]]]] 

```{r}

#visualize the colour centers for each image 

for (i in 1:length(rc_list)) {
  
  plotColorClusters(rc_list[[i]]$centers, rc_list[[i]]$sizes, 
                  plus = .5,
                  color_space = "sRGB", 
                  xlab = "red", ylab = "green", zlab = "blue", 
                  mar = c(3, 3, 2, 2),
                  main = "Histogram method of forming colour centers")
}

```

[[[[VISUALIZE II]]]]

```{r}

# plot a side-by-side comparison of an original image and its recolorized version, plus the color palette used for recoloring (I still don't understand why my original image is still so squashed)

for (i in 1:length(rc_list)) {
  
  plot(rc_list[[i]], 
       mar = c(3, 3, 2, 2), 
       main = "Histogram method of forming colour centers")
}

```



```{r}
# get a dataframe of all colors generated from recolorize2() in the recolorize objects in rc_list
all_palettes <- do.call(rbind, lapply(rc_list, function(i) i$centers))

# get a dataframe for cluster sizes (as a proportion of their original image):
all_sizes <- do.call(c, lapply(rc_list, function(i) i$sizes))

# plot colors using hclust_color() and form a grouping list first of 3 and 2. In the resulting lists, each element is a vector of which of the original colors should be clustered together. By default, hclust_color() calculates the Euclidean distance matrix between all provided color centers in CIE Lab color space

cluster_list <- hclust_color(all_palettes, n_final = 3)

cluster_list2 <- hclust_color(all_palettes, n_final = 2)

# the number indices for each colour blob here don't really mean anything - they're just colour indices used in the package? I think?
```

(next two code patches) Making a universal colour pallete to impose on to all of my images in a data set. Every pixel is assigned to the color it is closest to in some specified color space. I am doing this because this makes the most sense in terms of batch processing. This process involves extracting a specifeid number of colours from the colour cluster list generated from all of the specimen image, then mapping them onto every image, so that the color centers (and the subsequent colour maps) are the same across all of my images.

```{r}
# make an empty array for storing the new palette. ncol = 3 because 
elytra_palette2 <- matrix(NA, ncol = 3, nrow = length(cluster_list2))

# Using the list formed by hclust_colour() to - for every color in cluster_list...
for (i in 1:length(cluster_list2)) {
  
  # get the center indices
  idx <- cluster_list2[[i]]
  
  # get the average value for each channel, using cluster size to get a weighted average
  ctr <- apply(all_palettes, 2,
                 function(j) weighted.mean(j[idx], w = all_sizes[idx]))
  
  # store in the palette matrix
  elytra_palette2[i, ] <- ctr
}

# check that our colors seem reasonable
par(mar = rep(0, 4))
plotColorPalette(elytra_palette2)
```

```{r}

# use imposeColors() to map every image to the same set of two colors as earlier defined by hcluster_color(). Imposecolors takes an image and a set of color centers, and assigns each pixel to the most similar provided color as earlier defined. This is done for batch processing because it produces a set of images with identical colors. 
impose_list2 <- lapply(imgs, function(i) imposeColors(i, elytra_palette2, 
                                                     adjust_centers = FALSE, 
                                                     plotting = TRUE))
```

```{r}

# visualize palletes and recolorized object imaages 

layout(matrix(1:10, nrow = 2, ncol = 5))
par(mar = rep(0, 4))
for (i in impose_list2) { 
  plotImageArray(recoloredImage(i))
  plotColorPalette(i$centers, i$sizes, horiz = FALSE)
}
```

Converteach recolorize fit in back to a patternize format using recolorize_to_patternize() so that we can continue analyses in patternize

[[[[CONVERT BACK TO PATTERNIZE]]]]]
```{r}

# convert to patternize:
patternize_list2 <- lapply(impose_list2, recolorize_to_patternize)
patternize_test <- lapply(impose_list2, recolorize_to_patternize)

extent_list <- lapply(imageList_aligned, extent)

# set extents again:
for (i in 1:length(patternize_list2)) {
  for (j in 1:length(patternize_list2[[1]])) {
    raster::extent(patternize_list2[[i]][[j]]) <- extent_list[[i]]
  }
}
```

[[[[PatPCA_total]]]]
```{r}

#plot PCA using custom function 
setwd("C:/Users/USER/Documents/My School Documents/UofA/Lightroom")

source("patPCA_total.R")

test_pca <- patPCA_total(patternize_list2, quietly = FALSE)

```

[[[[Struggling to get legible data from the PCA. How do I read the rotations and loadings of the PCA plot? What variables are being used to assess variation in the PCA? shape? colour?]]]]

```{r}
patternize_merged <- vector("list", length(patternize_test))

for (i in 1:length(patternize_test)) {
    patternize_merged[i] <- do.call(merge, patternize_test[[i]])
}

IDlist <- c('CNC1151335','CNC1151393', 'CNC1860122','CNC1860123','CNC1860300')
```


[[[[[trying to run a regular patternize PCA]]]]

```{r}
# sum colour patterns?
summedtest <- sumRaster(patternize_test, IDlist, type = "k")
#summedtest2 <- sumRaster(patternize_test, IDlist)


# making theoretical populations from the five practice images 
pop1 <- c("CNC1151335","CNC1151393", "CNC1860300")
pop2 <- c("CNC1860122","CNC1860123")
popList <- list(pop1, pop2)
colList <- c("red", "blue")
symbolList <- c(16,17)


patPCA(patternize_merged, 
       popList,
       colList,
       symbolList = symbolList, 
       plot = TRUE,
       plotType = "points",
       plotChanges = TRUE,
       plotCartoon = TRUE,
       refShape = "target",
       refImage = imageList[["CNC1151335"]], 
       outline = mask,
       flipRaster = "y", 
       imageList = imageList,
       normalized = TRUE,
       cartoonID = "CNC1151335",
       cartoonFill = "black", 
       cartoonOrder = "under", 
       legendTitle = "Differences")

```

